---
title: "AFAM_Quick_Start"
author: "Sustainable Fisheries Group"
date: "March 24, 2017"
output: html_document
---

Simply copy and paste the following code into your R console. This will load the AFAM package onto your computer from an online file sharing service. Note that this step requires internet connection.

```{r setup, echo=TRUE,eval=FALSE}
## Specify packages that are required for AFAM app. Install them if they're not installed already
list.of.packages <- c("tidyverse", "gridExtra", "pander", "shiny", "DT","devtools")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages) > 0 ) install.packages(new.packages)

## Load packages required for AFAM
lapply(list.of.packages, require, character.only = TRUE)

## Install AFAM from Github, an online file-sharing service
install_github("SFG-UCSB/afamAppPackage")

## Load AFAM Package
library(afamAppPackage)

## Run AFAM App
runAFAM()
```

You now have the AFAM App Package installed on your computer! To use it, simply load the package, and run the app. You may cut and paste the following code into your R console. Note that once the package is installed during the step above, you no longer need internet connectivity to run the app!

```{r setup, echo=TRUE,eval=FALSE}
## Load AFAM Package
library(afamAppPackage)

## Run AFAM App
runAFAM()
```